---
title: "Final Project"
subtitle: "Data Science for Biologists, Spring 2021"
author: "Sam Dorofy"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# DON'T CHANGE THE NEXT LINE, except to add NEW options if you have any.
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo = TRUE, include = TRUE)

# Load all libraries you use HERE:
library(tidyverse)
library(dplyr)
library(modelr)
library(pROC)
```

<br><br>

## Introduction

```{r read-examine-data}
brain_tumors <- read_csv("dorofy_braintumor.csv")
glimpse(brain_tumors)
view(brain_tumors)
```

<br><br>

## Question 1

####Entropy in MRI scans is used to determine the size and location a possible brain tumor while Homogeneity refers to the balance of the magnetic field with or without a patient present.  What is the relationship between the Entropy and Homogeneity in the brain tumor MRI scans?

### Methods
```{examine mri data}
### Methods 
brain_tumors_nona <- brain_tumors %>% drop_na()

brain_tumors %>%
  ggplot(aes(x = Homogeneity, y = Entropy, fill = Target)) +
  geom_point() +
  scale_fill_gradient()) +
  labs(x = "Homogeneity",
       y = "Entropy",
       title = "Homogeneity and Entropy Values in the Presence of both Brain Tumors and no Brain Tumors")


### Answer

Replace this text with your answer to your question.

```
<br><br>

## Question 2

#### Of all of the patients in the study, how many have a brain tumor or no tumor?  Determine the number of each and exhibit these findings visually.  

```{Count of Tumor Presence}

### Methods
brain_tumors$Target<-ifelse(brain_tumors$Target==0, "yes","no") 
#'0' is no tumor and '1' is a tumor
brain_tumors %>%
  ggplot(aes(x = Target, fill = Target)) +
  geom_bar(color = "black") +
  scale_fill_manual(values = c("darkmagenta","pink")) +
  labs(x = "Presence of Brain Tumor",
       y = "Number of Patients",
       title = "How Many Patients Have a Brain Tumors Present?") 


### Answer

There are substantially more patients with a brain tumor as opposed to no brain tumor.
```
<br><br>

## Question 3

#### How many images have a mean above 10? Do these patients have a high Entropy value?

### Methods
```{}
brain_tumors$Target<-ifelse(brain_tumors$Target==0, "yes","no") 
brain_tumors %>%
filter(Mean>=10) %>%
count()
brain_tumors %>%
 ggplot(aes(x = Entropy, y = Mean, fill = Target)) +
  geom_point() +
  scale_fill_manual(values = c("darkmagenta","pink")) +
  labs(x = "Entropy Value",
       y = "Mean Above 10",
       title = "How Many Patients Have a Brain Tumors Present?") 






### Answer

## Replace this text with your answer to your question.


```
<br><br>

## Question 4

#### Is an increased Standard Deviation related to a high Homogeneity?

```{std dev and homogeneity relationship}

### Methods

### Answer

Replace this text with your answer to your question.
